### 2018-02-23 thesis/contents/discussions.tex

The segmentation of the subregions of the trigeminal pathway allowed for the possibility of measuring the respective diffusivities. Moreover, this method can extend to any scalar images that can be registered to the DWI space. These include specific tissue-metrics such as myelin, G-ratio, and PET images. This extended these methods to facilitate disease studies well beyond TN. We were able to divide the trigeminal nerve into 30 discrete cross-sections along its path, and observed detailed diffusivity changes as the nerve transitioned from pontine tissue to the cistern. The number of divisions chosen was arbitrary, with consideration that the number 30 could easily be divided into 3 sub-segments to represent the beginning, middle, and end of an elongated structure. In theory the number of divisions could be infinitely large, given that a centroid curve was identified. In practice the cross-sectional volume decreased with more divisions, and number of points that were sampled by each cross-sectional volume needed to be considered. Determining a common centroid curve from many subjects was computationally non-trivial, especially when the streamline lengths were highly variable due to the noise in the cistern CSF. White matter pathways are highly complex, with possible multi-directional forks, bends, and convergent pathways across a wide range of scales, some examples include the forniceal crura, zona incerta, and optic chaism. A generalized algorithm that can reliably calculate a centroid for these structure is still a topic of research. Turns out that the merged tractography strategy simplified this task by allowed a single curve to be defined manually in the template tractography space. This greatly reduced the amount of labour for researchers, as centroids no longer need to be defined repeatedly for each subject. Manually curve definition by experts can leverage known techniques in computer graphics to produce centroid for very complex structures, thereby extending SAGIT for obtaining along-the-track measure for other highly variable streamline bundles.

@@

I) Keep pushing the limit of anatomical delineations towards other areas of the trigeminal system. This include the spinal trigeminal pathways, which include the spinal trigeminal ganglion and its sub-gangions, and its medullary decussations. Re-examine TPT tracts, both to establish its diffusivity baseline, and to confirm its asymmetry. Image the medial pain pathways, and descending projections to the PAG. These tasks are currently limited by the resolution of availabile DWI images. Immediate gains may be obtained by using SAGIT to delineate these structure on the Human Connectome Project (HCP) dataset. Additionally, investigation into 7T DWI images will be necessary.

>>>>

The segmentation of the subregions of the trigeminal pathway has allowed for the possibility of measuring the respective diffusivities. Moreover, this method can extend to any scalar images that can be registered to the DWI space. These include specific tissue-metrics such as myelin, G-ratio, and PET images. This methodology is therefore highly applicable and not specific to the study of TN. We divided the trigeminal nerve into 30 discrete cross-sections along its path, and observed detailed diffusivity changes as the nerve transitioned from pontine tissue to the cistern. The number of divisions chosen was arbitrary, with consideration that the number 30 could easily be divided into 3 sub-segments to represent the beginning, middle, and end of an elongated structure. In theory the number of divisions could be infinitely large, given that a centroid curve was identified. In practice the cross-sectional volume decreased with more divisions, and number of points that were sampled by each cross-sectional volume needed to be considered. Determining a common centroid curve from many subjects was computationally non-trivial, especially when the streamline lengths were highly variable due to the noise in the cistern CSF. White matter pathways are highly complex, with possible multi-directional forks, bends, and convergent pathways across a wide range of scales, some examples include the forniceal crura, zona incerta, and optic chaism. A generalized algorithm that can reliably calculate a centroid for these structure is still a topic of research. Turns out that the merged tractography strategy simplified this task by allowed a single curve to be defined manually in the template tractography space. This greatly reduced the amount of labour for researchers, as centroids no longer need to be defined repeatedly for each subject. Manually curve definition by experts can leverage known techniques in computer graphics to produce centroid for very complex structures, thereby extending SAGIT for obtaining along-the-track measure for other highly variable streamline bundles.

@@

I) Keep pushing the limit of anatomical delineations towards other areas of the trigeminal system. This include the spinal trigeminal pathways, which include the spinal trigeminal ganglion and its sub-ganglions, and its medullary decussations. Re-examine TPT tracts, both to establish its diffusivity baseline, and to confirm its asymmetry. Image the medial pain pathways, and descending projections to the PAG. These tasks are currently limited by the resolution of availabile DWI images. Immediate gains may be obtained by using SAGIT to delineate these structure on the Human Connectome Project (HCP) dataset. Additionally, investigation into 7T DWI images will be necessary.
